### 2018-10-10 rebuttal/response.txt
copied over files from bitbucket/akaposi/elims

Answers to reviewer A:

- The function space with small domain can be indeed seen as an
  operator which adds parameters to constructors.

- We don't use impredicativity in the paper.

Answers to reviewer B:

- We use the metatheoretic assumptions which are stated in section 2:
  UIP, function extensionality, and the QIIT of signatures.

- On well-behavedness of signatures: the definition of the syntax
  follows Altenkirch & Kaposi (POPL 2016), and it is based on CwF-s,
  which is a standard categorical presentation of the syntax. We add the
  non-standard function space with metatheoretic
  domain. Also, the syntax is an equational theory
  and the rules should not be read as rewriting rules, and equality
  derivations have to be explicitly constructed in the metatheory.

- The equivalence of elimination and initiality is a theorem. It follows
  from the CwF-K-Eq model, which is partially formalized
  in Agda.

- Delta^A is a large Sigma-type in the metatheory.

- Indeed, for instance, more terms are typable when the metatheory has
  function extensionality.  Consider the following context:

    (A    : Pi^ Nat \n -> U,
     con1 : Pi^ Nat \n -> A @^ n,
     con2 : Pi^ Nat \n -> A @^ (n + 0))

  In this context, con2 has type (Pi^ Nat \n -> A @^ n) only if we
  have funext in the metatheory. That's because metatheoretic
  functions appear in Pi^. This situation does not arise if we only
  have closed QIIT signatures.

Answers to reviewer C:

1. We will rename U to iota in the example. We've found that this is
   the smallest example with a meaningful equality constructor.

2. Yes, it quantifies over gamma implicitly. We will clarify this and
   refer to section 2 where this is explained.

3. We will clarify.

4. We include Top here to signal that ^M is to be computed entry-wise
   from signatures.

6. We improve on the FoSSaCS 2018 paper in the following ways. We will
   make this clear in the paper.

    - Construction of initial algebras.
    - Showing that being able to specify signatures is already enough to
      construct initial algebras.
    - Computing notions of displayed algebras and sections strictly, not
      just up to isomorphism.  Kaposi & Kovács FSCD 2018 demoed this
      with a computer program which outputs types of eliminators from
      input signatures. FoSSaCS 2018 used finite completeness as a
      property to talk about induction, while we use explicit CwF
      structure for the same purpose.

   We haven't investigated the relation between the semantic
   conditions in FoSSaCS 2018 and our syntax. This could be future
   work.

7. We will refer to section 2 before section 1.1.

8. We skip lambda only because we know of no QIIT examples where it is
   required.

9. Apart from section 1, we exclusively use de Bruijn indices in the
   theory of signatures, names are only for metatheoretic variables.

10. "2" denotes a de Bruijn variable.

11, 12: we'll improve the presentation.

Scaling to infinitary QIITs: we are confident that the solution
mentioned in section 8 works. A different potential solution is to
recover set-level univalence in models, by switching to a
groupoid-truncated syntax (extended with some coherence laws).